What do incoming university students believe about open science practices in

### 2021-03-24 analysis/Manuscript.Rmd
tidy up means table

---
title: What do incoming university students believe about open science practices in
  psychology?
author:
- name: Jennifer L. Beaudry
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: jbeaudry@swin.edu.au
  role: null
- name: Michael C. Philipp
  affiliation: '2'
  role: null
- name: Matt N. Williams
  affiliation: '2'
  role: null
shorttitle: Psychology students' beliefs about open science
output:
  papaja::apa6_word: default
  papaja::apa6_pdf: default
authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  <!-- https://tinyurl.com/ybremelq -->
keywords: open science, psychology, teaching, reproducibility, replication
wordcount: X
bibliography: ['r-references.bib','Undergrads & Open Science (public).bib']
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: yes
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Swinburne University of Technology
- id: '2'
  institution: Massey University
---
<!-- I've only partially entered authorship deets for the moment and set it to mask authorship info when knitting
It's set to default to knit to .docx, but can also knit to pdf (if you have a tex distro installed)
The pdf looks nicer and might be good for a preprint but .docx probably easiest to use for journal submissions.
To knit to RMarkdown you will probably need the papaja package installed, and to knit to .pdf you will need a tex distro (e.g., TinyTex).
However, it is possible to run all the R code (without knitting to .docx or .pdf) by clicking Run > Run All Chunks Below.
This could be useful in future for readers interested in reproducing results but who don't have or wish to install 
papaja and its many dependencies.
-->

@@

```{r setup, include = FALSE}
library(papaja)
library(here)
library(tidyverse)

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

The last decade has seen unprecedented change in methodological and reporting practices in psychology. These changes were partly precipitated by what is popularly known as the "replication crisis": The discovery that close replications of published psychological studies are often unable to replicate the original findings[@kleinInvestigatingVariationReplicability2014; @kleinManyLabsInvestigating2018]. These apparent problems with replication have lead to a variety of potential solutions to make research practices more transparent, including more frequent publication of replication studies [see @brandtReplicationRecipeWhat2014], more thorough reporting of methods and results [@simmons21WordSolution2012], open sharing of data [see @meyerPracticalTipsEthical2018], and preregistration of data collection and analysis plans [@nosekPreregistrationRevolution2018]. These calls for more transparent research practices prompted the discipline to reflect on the norms and beliefs that underpin its practice of research. Many studies have polled researchers' understandings of and adherence to open science norms, beliefs, and practices (e.g., @baker500ScientistsLift2016; @openscienceworkinggroup2017SurveyOpen2018, @harrisUseReproducibleResearch2018). In turn, initiatives have been directed toward training undergraduate students in the adoption of an open science ethos [e.g., @chopikHowWhetherTeach2018; @graheHarnessingUndiscoveredResource2012; @jekelHowTeachOpen2020; @schonbrodtTrainingStudentsOpen2019].  These initiatives help engrain open science norms and change attitudes about research practices, but we know little about what these students know or believe about open science research practices prior to entering the university classroom.

The most comprehensive set of principles for how science *ought* to be practiced are Merton's [@mertonNoteScienceDemocracy1942] norms of science. Scientists [@andersonNormativeDissonanceScience2007] and graduate students [@andersonGraduateStudentExperience1994] alike have historically endorsed the normative value of Merton's principles - to at least some extent. Furthermore, many of the recent practices designed to make science more transparent and open reflect Mertonian norms. For example, the practice of sharing open source software directly corresponds to the Mertonian norm of *communism*: Scientists should have common ownership of scientific goods. Similarly, the practices of sharing of preprints for open peer review and open data for checking of reproducibility corresponds to the norm of communality, but also to that of *organised skepticism*: Scientific claims should be subjected to critical scrutiny. Preregistration can be connected to the norm of *disinterestedness*: By making (and preregistering) decisions about how to analyse data before results are produced, a researcher can limit the degree to which the substantive results produced by different analytic strategies affect their decisions regarding which analyses to report. In this sense, the ongoing reform in psychological research can partly be understood as simultaneously a set of new practices and a re-affirmation of old norms.

@@

First, unlike some areas of psychology covered in undergraduate courses (e.g., models of working memory, or the internal workings of human senses), the replication crisis is frequently discussed in mainstream and social media [e.g., @yongPsychologyReplicationCrisis2016; @yongPsychologyReplicationCrisis2018; @ogradyPsychologyReplicationCrisis2020]. Many students may plausibly have some knowledge about replicability issues in psychology which they obtained prior to (or independently of) their formal studies. Teaching methods should thus be informed by some understanding of what students' typical pre-existing levels of knowledge are.

<!--Second, when psychology students are beginning their university studies, they are often learning for the first time about how and why it is useful to apply scientific methods to studying human behaviour (rather than only relying on alternative sources of knowledge such as intuition or anecdote or authority). Could an over-emphasis on problems with replicability leave such students unconvinced that scientific methods for studying human behaviour are valuable *at all*, leaving them to favour even less credible alternative sources of knowledge? It is therefore important to determine what levels of trust in psychological research (and its replicability) are prevalent in incoming undergraduate students. In their examination of the usefulness of a one-hour lecture about the replication crisis, @chopikHowWhetherTeach2018 found that undergraduate students trusted the results of studies by psychologists less after the lecture than they did before it, although the effect size was fairly small (d = -.36), and mean trust levels remained fairly high after the lecture (M = 4.94 on a scale of 1 to 7).

#I've commented this paragraph out because on reflection I'm not sure it's a very strong argument, and the potential effect of teaching about replicability on levels of trust in psych science isn't really something we can directly address in this study.
However, it was one of the only places where we really talked about the Chopik et al. 2018 paper, which probably needs more coverage.

-->

Second, anecdotal evidence suggests that students being taught about open science practices and reforms to improve reproducibility (such as open sharing of data and analysis code) are often surprised that this is not *already* standard practice [e.g., @duyguuygun-tuncLastNightWas2021]. In other words, these is a possibility that students' naive conception of how science works may in some respects be closer to that embodied in recent reforms rather than "business as usual" practices. Indeed, surveys of scientists' subscriptions to Mertonian norms have often found that while most scientists endorse Mertonian norms such as communality, universalism and disinterestedness, this is by no means the case for all scientists [@andersonExtendingMertonianNorms2010]. Furthermore, scientists typically perceive the behaviour of other scientists as being less consistent with these norms than their own [@andersonNormativeDissonanceScience2007]. It is thus possible that, for some students, teaching about reform might in some cases involve reinforcing "naive" assumptions. In fact, although beyond the scope of this article, it is *possible* that incoming undergraduate students more strongly endorse open science norms than do academics.

@@

Third, undergraduate psychology students themselves represent one of the most important audiences for psychological research: The number of undergraduate students enrolled in psychology courses in any given year dwarfs the number of academics working in psychology.<!--I assume this is true but cannot find any good numbers to pack it up! --> Furthermore, by serving as participants for course credit or cash payments [@bowenCashCreditCompensation2017], undergraduate psychology students provide the data underlying a substantial portion of psychological research. For example, @thalmayerNeglected95Revisited2021 found that 42% of studies reported in the *Journal of Personality and Social Psychology* in the period 2014 to 2018 were based on undergraduate student samples. As such, undergraduate students are both key consumers of psychological research as well as much-needed suppliers of the data used in psychological research. Therefore, the *preferences* of these students with respect to reproducibility and open science practices are themselves important as a phenomenon of interest. If, for example, undergraduate psychology students have a strong preference that journal articles are freely available, and that the data underlying claims is widely shared, these preference on the part of some of the key stakeholders in psychological research should have some bearing on our willingness to engage in these practices.

<!-- I'm not sure about the argument in the above paragraph - might be somewhat of a stretch?
9Mar: On reflection I think it's kinda an ok argument, albeit we need more evidence for the claims in it. 
I almost want to phrase it as saying that "Undergraduate students are both our most important customers and our most important suppliers of data"...
-Matt
-->

These considerations suggest that it is important that teachers wishing to inform undergraduate psychology students about the replication crisis and open science practices have some understanding of what such students know and believe about these topics already. Currently, relatively little is know about undergraduate students' beliefs about replicability and open science practices. A study of by @chopikHowWhetherTeach2018 (focused on the effect of a one-hour lecture about open science) found that students tended to endorse pro-replicability/open science beliefs after the lecture (e.g., 92.7% thought it was "very" or "extremely" important that researchers "make data publicly available so that results can be verified by other researchers"). However, this study was obviously not

@@

# once qual responses are coded, I will need to return here to 
 # include details about which univrsity the participant is enrolled in
 # name of qualification
 # major [how many are majoring in psyc; NAs; will decide later?]

@@

### Open Science Norms and Counternorms

Our approach to measuring norms and counternorms focussed on participants' evaluation of ten specific open science practices that reflected open science norms rather than trying to measure endorsement of the more abstract Mertonian norms (cf. Anderson et al., 2007). 
The open science practices assessed in our survey included: critical thinking, preregistration, registered reports, p-hacking, hypothesising after the results are known (HARKing), providing sufficent information of replicability, preprints, open materials, open data, and open access publishing. Each practice was elucidated by a pair of items (pro-open science; counter-open science) in the context of a common scenario, "Imagine that Deborah is a psychology researcher who has designed a study to test a specific hypothesis." There were 20 scenarios in total.
For example, the practice of posting article preprints was evaluated with the pro-open science item, "Deborah should post a manuscript describing the Findings of her study openly online as soon as it is complete, even if the manuscript has not yet been checked by experts (peer reviewed) and accepted for publication in a journal." The counter-open preprint item read, "Deborah should not post a manuscript describing the findings of her study online until after it has been checked by experts (peer reviewed) and accepted for publication in a journal." Participants rated their agreement with each prescriptive statement on a 1 (strongly disagree) to 5 (strongly agree) scale.

### Norms in Practice

We also sought to measure participants' normative beliefs about how psychological science was actually conducted by researchers around the world. To that end we asked participants, "When you consider all of the psychology studies conducted globally each year, what percentage of them would have the following characteristics?" The ten characteristics explicitly described each of the ten open science practices named above. For example, "A manuscript describing the findings of the study is posted openly online as soon as it is complete, even if the manuscript has not yet been checked by experts (peer reviewed) and accepted for publication in a journal." Participants were given a slider bar ranging from 0% to 100% to record their estimate.

### Beliefs About Replications

Participants' beliefs about the replicabilty of results in psychological science were indexed by two questions. The first asked participants to "Imagine that a set of researchers selected 100 published studies about psychology, and repeat each study again, exactly how it was described in the original report but with a new set of participants. We call repeating a study like this a “replication study”. At a guess, how many of these replication studies do you think would produce the same conclusions as the original studies?” Estimates were recorded using a slider bar ranging from 0% to 100%. The second question asked participants whether they had heard of the replication crisis. For those who said they had, two open-ended

@@

@@

# create a `diff` variable that is the counternorm subtracted from the norm
norm_cnorm <- norm_cnorm %>% 
  mutate (diff = (`norm rating` - `cnorm rating`)) %>% 
      dplyr::select(-(`id`)) # remove this now that I've confirmed it works

norm_means <- norm_cnorm %>% 
    dplyr::group_by(practice) %>%  
    summarise(norm_mean = mean(`norm rating`, na.rm = TRUE), 
              norm_sd = sd(`norm rating`, na.rm = TRUE), 
              cnorm_mean = mean(`cnorm rating`, na.rm = TRUE), 
              cnorm_sd = sd(`cnorm rating`, na.rm = TRUE), 
              diff_mean = mean(`diff`, na.rm = TRUE), 
              diff_sd = sd(`diff`, na.rm = TRUE)) %>% 
  ungroup()

# breadcrumb: I need to use papaja tables....
# pull means & sd into the same column with (sd) in parentheses
# headings should then be norm, counternorm, difference

knitr::kable(norm_means,
             caption = "Means (sd) for norms, counternorms, and the difference between the two for each of the practices",
             digits = 2,
             align = c('lcccccc'))

>>>>

---
title: What do incoming university students believe about open science practices in
  psychology?
author:
- name: Jennifer L. Beaudry
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: jbeaudry@swin.edu.au
  role: null
- name: Michael C. Philipp
  affiliation: '2'
  role: null
- name: Matt N. Williams
  affiliation: '2'
  role: null
shorttitle: Psychology students' beliefs about open science
output:
  papaja::apa6_pdf: default
  papaja::apa6_word: default
authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  <!-- https://tinyurl.com/ybremelq -->
keywords: open science, psychology, teaching, reproducibility, replication
wordcount: X
bibliography: ['r-references.bib','Undergrads & Open Science (public).bib']
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: yes
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Swinburne University of Technology
- id: '2'
  institution: Massey University
---
<!-- I've only partially entered authorship deets for the moment and set it to mask authorship info when knitting
It's set to default to knit to .docx, but can also knit to pdf (if you have a tex distro installed)
The pdf looks nicer and might be good for a preprint but .docx probably easiest to use for journal submissions.
To knit to RMarkdown you will probably need the papaja package installed, and to knit to .pdf you will need a tex distro (e.g., TinyTex).
However, it is possible to run all the R code (without knitting to .docx or .pdf) by clicking Run > Run All Chunks Below.
This could be useful in future for readers interested in reproducing results but who don't have or wish to install 
papaja and its many dependencies.
-->

@@

```{r setup, include = FALSE}
library(papaja)
library(here)
library(tidyverse)
library(finalfit) # to keep trailing zeros

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed,
  results = 'as.is')
```

@@

The last decade has seen unprecedented change in methodological and reporting practices in psychology. These changes were partly precipitated by what is popularly known as the "replication crisis": The discovery that close replications of published psychological studies are often unable to replicate the original findings[@kleinInvestigatingVariationReplicability2014; @kleinManyLabsInvestigating2018]. These apparent problems with replication have lead to a variety of potential solutions to make research practices more transparent, including more frequent publication of replication studies [see @brandtReplicationRecipeWhat2014], more thorough reporting of methods and results [@simmons21WordSolution2012], open sharing of data [see @meyerPracticalTipsEthical2018], and preregistration of data collection and analysis plans [@nosekPreregistrationRevolution2018]. These calls for more transparent research practices prompted the discipline to reflect on the norms and beliefs that underpin its practice of research. Many studies have polled researchers' understandings of and adherence to open science norms, beliefs, and practices [e.g., @baker500ScientistsLift2016; @openscienceworkinggroup2017SurveyOpen2018, @harrisUseReproducibleResearch2018). In turn, initiatives have been directed toward training undergraduate students in the adoption of an open science ethos [e.g., @chopikHowWhetherTeach2018; @graheHarnessingUndiscoveredResource2012, @jekelHowTeachOpen2020, @schonbrodtTrainingStudentsOpen2019].  These initiatives help engrain open science norms and change attitudes about research practices, but we know little about what these students know or believe about open science research practices prior to entering the university classroom.

The most comprehensive set of principles for how science *ought* to be practiced are Merton's (1942) norms of science. Scientists (Anderson et al., 2000) and graduate students (Anderson & Louis, 1994) alike have historically endorsed the normative value of Merton's principles. Furthermore, many of the recent practices designed to make science more transparent and open reflect Mertonian norms. For example, the practice of sharing open source software directly corresponds to the Mertonian norm of *communism*: Scientists should have common ownership of scientific goods. Similarly, the practices of sharing of preprints for open peer review and open data for checking of reproducibility corresponds to the norm of communality, but also to that of *organised skepticism*: Scientific claims should be subjected to critical scrutiny. Preregistration can be connected to the norm of *disinterestedness*: By making (and preregistering) decisions about how to analyse data before results are produced, a researcher can limit the degree to which the substantive results produced by different analytic strategies affect their decisions regarding which analyses to report. In this sense, the ongoing reform in psychological research can partly be understood as simultaneously a set of new practices and a re-affirmation of old norms.

@@

First, unlike some areas of psychology covered in undergraduate courses (e.g., models of working memory, or the internal workings of human senses), the replication crisis is frequently discussed in mainstream and social media [e.g., @yongPsychologyReplicationCrisis2016; @yongPsychologyReplicationCrisis2018; @ogradyPsychologyReplicationCrisis2020]. Many students may plausibly have some knowledge about these issues obtained prior to (or independently of) their formal studies. Teaching methods should thus be informed by some understanding of what students' pre-existing levels of knowledge are.

Second, when psychology students are beginning their university studies, they are often learning for the first time about how and why it is useful to apply scientific methods to studying human behaviour (rather than only relying on alternative sources of knowledge such as intuition or anecdote or authority). Could an over-emphasis on problems with replicability leave such students unconvinced that scientific methods for studying human behaviour are valuable *at all*, leaving them to favour even less credible alternative sources of knowledge? It is therefore important to determine what levels of trust in psychological research (and its replicability) are prevalent in incoming undergraduate students. In their examination of the usefulness of a one-hour lecture about the replication crisis, @chopikHowWhetherTeach2018 found that undergraduate students trusted the results of studies by psychologists less after the lecture than they did before it, although the effect size was fairly small (d = -.36), and mean trust levels remained fairly high after the lecture (M = 4.94 on a scale of 1 to 7).

Third, anecdotal evidence suggests that students being taught about open science practices and reforms to improve reproducibility (such as open sharing of data and analysis code) are often surprised that this is not *already* standard practice. In other words, students' naive conception of how science works may in some ways be closer to that embodied in recent reforms rather than "business as usual" practices. Indeed, surveys of scientists' subscriptions to Mertonian norms have often found that while most scientists endorse Mertonian norms such as communality, universalism and disinterestedness, this is by no means the case for all scientists. Furthermore, they typically perceive the behaviour of other scientists as being less consistent with these norms than their own [@andersonNormativeDissonanceScience2007]. It is thus possible that, for some students, teaching about reform may be less a matter of conveying new information and more one of reinforcing "naive" assumptions. In fact, although beyond the scope of this article, it is *possible* that incoming undergraduate students more strongly endorse open science norms than do academics.

@@

These considerations suggest that it is important that teachers wishing to inform undergraduate psychology students about the replication crisis and open science practices have some understanding of what such students actually know and believe about these topics already. In addition, undergraduate psychology students themselves represent one of the most important audiences for psychological research: The number of undergraduate students enrolled in psychology courses in any given year dwarfs the number of academics working in psychology. As such, the *preferences* of these students with respect to reproducibility and open science practices are themselves important as a phenomenon of interest. If, for example, undergraduate psychology students have a strong preference that journal articles are freely available to members of the public, this preference on the part of some of the consumers of the knowledge we produce should have some bearing on our choices in relation to sharing of manuscripts. This is especially the case given that undergraduate students are directly or indirectly responsible for much of the funding which allows universities to operate and research to be conducted. In fact, in serving as participants for course credit in many universities, undergraduate psychology students also provide the data underlying much psychological research.

<!-- I'm not sure about the argument in the above paragraph - might be somewhat of a stretch? -->

@@

# [breadcrumbs] 
 # once qual responses are coded, I will need to return here to 
 # include details about which univrsity the participant is enrolled in
 # name of qualification
 # major [how many are majoring in psyc; NAs; will decide later?]

@@

@@

In line with our preregisterd analysis plan, we cacluated the difference in means
between responses to each norm item and its accompanying counternorm, along with 
95% confidence intervals. For this analysis, we removed three cases with missing values. 
Table XX displays the means and standard deviations for each norm, counternorm, 
and the difference values.

@@

# create a `diff` variable that is the counternorm subtracted from the norm
norm_cnorm <- norm_cnorm %>% 
  mutate (diff = (`norm rating` - `cnorm rating`)) %>% 
    drop_na() %>% # drop the three NA values
      dplyr::select(-(`id`)) # remove this now that I've confirmed it works

# get the n values for each of the practices
prac_counts <- norm_cnorm %>% 
    dplyr::group_by(practice) %>%  
    count()

# calculate means
prac_ms_sds <- norm_cnorm %>% 
    dplyr::group_by(practice) %>%  
    summarise(norm_mean = mean(`norm rating`), 
              norm_sd = sd(`norm rating`),
              cnorm_mean = mean(`cnorm rating`), 
              cnorm_sd = sd(`cnorm rating`),
              diff_mean = mean(`diff`), 
              diff_sd = sd(`diff`)) %>% 
  ungroup()

# join the n values with the means
prac_means <- prac_counts %>% inner_join(prac_ms_sds)

# calculate 95% CIs
error <- qnorm(0.975) * prac_means$diff_sd/sqrt(prac_means$n)
prac_means$lower <- prac_means$diff_mean - error
prac_means$upper <- prac_means$diff_mean + error

# combine means & SDs
prac_means$Norm <- paste0(round_tidy(prac_means$norm_mean,2),
                          " (", 
                          round_tidy(prac_means$norm_sd,2), 
                          ")")
prac_means$Counternorm <- paste0(round_tidy(prac_means$cnorm_mean,2), 
                            " (", 
                            round_tidy(prac_means$cnorm_sd,2), 
                            ")")
prac_means$Difference <- paste0(round_tidy(prac_means$diff_mean,2), 
                            " (", 
                            round_tidy(prac_means$diff_sd,2),
                            ")", 
                            " [",
                            round_tidy(prac_means$lower,2),
                            ", ", 
                            round_tidy(prac_means$upper,2),
                            "]")

# clean it up into a new dataframe
practice_means <- prac_means %>% 
  select(practice,Norm, Counternorm, Difference)

# breadcrumb: I need to use papaja tables....
# pull means & sd into the same column with (sd) in parentheses
# first level headings should then be norm, counternorm, difference

knitr::kable(practice_means,
             caption = "Means (sd) for norms, counternorms, and the difference between the two for each of the practices",
             digits = 2,
             align = c('lccc'))
